### 2020-01-22 probabilisticProgrammingMartingales.tex
Added some thoughts on further directions / next steps.

\lo{Is the induced $\sigma$-algebra the Borel algebra?
The basis elements of the topology of $S$ have the form $\prod_{i \in I} U_i \times \mathbb{R}^\omega$ where $I$ ranges over finite sets and each $U_i$ is open in $\mathbb{R}$.}

The one-step reduction is given by the function $\red : \Lambda \times S \to \Lambda \times S$ where
\begin{equation}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s) & \text{if } M = E[R], R \to N \text{ and } R \neq \sample \\
        (\pi_h(s),\pi_t(s)) & \text{if } M =  \sample \\
        (M,s) & \text{if } M \text{ normal form}
    \end{array} \right .
\end{equation}
The limit $\red^\infty$ can then be defined as a partial function as $\lim_{n \to \infty} \red^n(M,s)$ whenever that sequence becomes constant by reaching a normal form.

\begin{comment} I'm not sure whether this section will be needed any more.
\section{Recursive steps}
Define a reduction relation on PPCF terms that excludes the $\Y$ combinator (``nonrecursive reduction''), i.e. $E[M] \dashrightarrow E[N] \text{ if } M \to N$ for any redex $M$ not of the form $\Y M'$.

This reduction relation is strongly normalizing because of the simple type system, so for any term $M$ there is some bound $b$ such that nonrecursive reduction of $M$ must terminate in at most $b$ steps. Define $\red_{nr} : \Lambda \times S \to \Lambda \times S$ like $\red(M,s)$, but using nonrecursive reduction instead of reduction, and let $\red_Y(M,s) = \red^{b_M}_{nr}(M,s)$. The sequence $(r_\Lambda(M,s,n), r_S(M,s,n)) = r(M,s,n) = \left(\red_Y \circ \red \right)^n(\red_Y(M,s))$ is then a subsequence of $\left( \red^n(M,s) \right)_n$ consisting of only terms which are either normal form or of the form $E[\Y N]$ for some $N$.
\end{comment}

@@

\lo{Further directions - suggestions to come.}

>>>>

\lo{Is the induced $\sigma$-algebra the Borel algebra?
The basis elements of the topology of $S$ have the form $\big(\prod_{i \in I} U_i\big) \times \mathbb{R}^\omega$ where $I$ ranges over finite sets and each $U_i$ is open in $\mathbb{R}$.}

The one-step reduction is given by the function $\red : \Lambda \times S \to \Lambda \times S$ where
\begin{equation}
\red(M,s) = \left\{
    \begin{array}{ll}
        (E[N],s) & \text{if } M = E[R], R \to N \text{ and } R \neq \sample \\
        (\pi_h(s),\pi_t(s)) & \text{if } M =  \sample \\
        (M,s) & \text{if } M \text{ normal form}
    \end{array} \right .
\end{equation}
\lo{Your definition $\red$ is incomplete: it is missing the definition of $R \to N$.}

The limit $\red^\infty$ can then be defined as a partial function as $\lim_{n \to \infty} \red^n(M,s)$ whenever that sequence becomes constant by reaching a normal form.

%\begin{comment} 
\akr{\section{Recursive steps}
[I'm not sure whether this section will be needed any more.]

Define a reduction relation on PPCF terms that excludes the $\Y$ combinator (``nonrecursive reduction''), i.e. $E[M] \dashrightarrow E[N] \text{ if } M \to N$ for any redex $M$ not of the form $\Y M'$.

This reduction relation is strongly normalizing because of the simple type system, so for any term $M$ there is some bound $b$ such that nonrecursive reduction of $M$ must terminate in at most $b$ steps. Define $\red_{nr} : \Lambda \times S \to \Lambda \times S$ like $\red(M,s)$, but using nonrecursive reduction instead of reduction, and let $\red_Y(M,s) = \red^{b_M}_{nr}(M,s)$. The sequence $(r_\Lambda(M,s,n), r_S(M,s,n)) = r(M,s,n) = \left(\red_Y \circ \red \right)^n(\red_Y(M,s))$ is then a subsequence of $\left( \red^n(M,s) \right)_n$ consisting of only terms which are either normal form or of the form $E[\Y N]$ for some $N$.
}
%\end{comment}

@@

\lo{\textbf{Further directions}

\smallskip

An obvious next step is to extend the result to the $\mathsf{score}$ construct.

The following are highly topical, and could form the basis of an interesting and novel DPhil thesis.
\begin{enumerate}
\item Devise methods for proving (positive) a.s.~termination.

- For example, develop a type system satisfying the property: if a term is typable then it is (positively) a.s.~terminating.
See~\cite{DBLP:conf/ppdp/BreuvartL18,DBLP:conf/esop/LagoG17}.

- Another approach is to develop algorithms that synthesise ranking supermartingales, following, for example, \cite{DBLP:journals/pacmpl/AgrawalC018}.

\item Develop principles (e.g.~in the form of ``proof rules'') for reasoning about (positively) a.s.~termination, in the style of \cite{DBLP:journals/pacmpl/McIverMKK18}.

\item Design algorithms that synthesise probabilistic invariants (\emph{qua} martingales), \`a la \cite{SchreuderO19}; see also \cite{DBLP:journals/pacmpl/HarkKGK20}.

\end{enumerate}}
